---
title: "p8105_hw6"
output: github_document
author: Carolyn Fish (csf2135)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

#### Problem 1

The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.

* Create a city_state variable (e.g. “Baltimore, MD”) [x]
* a binary variable indicating whether the homicide is solved. [x]
* Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. [x] 
* Also omit Tulsa, AL – this is a data entry mistake. [x]
* For this problem, limit your analysis those for whom victim_race is white or black. [x]
* Be sure that victim_age is numeric. [x]
```{r, message=FALSE}
homicide_df = 
  read_csv("homicide-data.csv", 
           na = c("", "NA", "Unknown"), 
            col_types = cols(
                         victim_age = col_double()
                       ))  |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolution = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest"        ~ "unsolved",
      disposition == "Closed by arrest"      ~ "solved"
    )
  ) |> 
filter(
  !(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO")),
  victim_race %in% c("Black", "White")
)
```

* For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. [x]
* Save the output of glm as an R object; [x]
* apply the broom::tidy to this object; [x]
* and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed. [x]
```{r, message = FALSE, warning= FALSE}
# Recode outcome = unsolved (0), and outcome = solved (1)
# Recode race = white as reference (0), and race = black (1)
# Recode sex = female as reference (0), and sex = male (1)

homicide_df = homicide_df |> 
  filter(
    !is.na(victim_sex) & !is.na(victim_age) & !is.na(victim_race) & !is.na(resolution)
  ) |> 
  mutate(
    outcome = case_match(
      resolution, 
     "solved" ~ 1, 
     "unsolved" ~ 0 
    ), 
    race = case_match(
      victim_race, 
      "Black" ~ 1, 
      "White" ~ 0
    ), 
    sex = case_match(
      victim_sex, 
      "Female" ~ 0, 
      "Male" ~ 1
    )) |> 
  select(
    city_state, outcome, race, sex, victim_age
  )
```

```{r, message = FALSE, warning=FALSE}
# fit logistic regression for resolution in Baltimore, MD
fit = glm(outcome ~ victim_age + sex + race, data = homicide_df |> filter(city_state == "Baltimore, MD") , 
          family = "binomial")

#create table of terms with confidence intervals
tidy_summary = broom::tidy(fit)

conf_intervals = confint(fit)

result_baltimore = cbind(tidy_summary, conf_intervals) 

result_baltimore |> 
  mutate(
    adj_OR = exp(estimate), 
    `lower 95% CI` = exp(`2.5 %`), 
    `upper 95% CI` = exp(`97.5 %`),
  ) |> 
  select(
    adj_OR, `lower 95% CI`, `upper 95% CI`
  ) |> 
  kableExtra::kable(digits = 2 )
```

**Interpretation:** The odds of a homicide case in Baltimore, MD being solved when victims were male is 0.43 times that of cases where victims were female, adjusting for the victim's age and race. At the 95% level of significance, a range of reasonable estimates for the odds ratio are between 0.32 and 0.56. The null hypothesis of an adjusted OR of 1 can be rejected. 

* Now run glm for each of the cities in your dataset, and 
* extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. 
* Do this within a “tidy” pipeline, 
  * making use of purrr::map, 
  * list columns, and 
  * unnest as necessary to create a dataframe with estimated ORs and CIs for each city.
```{r}
# create a new function to calculate adjusted OR and 95% CI
fit_and_extract <- function(df) {
  fit = glm(outcome ~ victim_age + sex + race, data = df, family = "binomial")
  
  tidy_summary = broom::tidy(fit)
  
  conf_intervals = confint(fit)
  
  result = cbind(tidy_summary, conf_intervals)
  
  result  |> 
    filter(term == "sex") |> 
    mutate(
      adj_OR = exp(estimate),
      `lower 95% CI` = exp(`2.5 %`),
      `upper 95% CI` = exp(`97.5 %`)
    )  |> 
    select(adj_OR, `lower 95% CI`, `upper 95% CI`)
}
```
  
```{r, message = FALSE, warning= FALSE}
results_all = homicide_df |> 
  group_by(city_state) |> 
  nest(data = c(outcome, victim_age, sex, race)) |> 
  mutate(
    results = purrr::map(data, fit_and_extract)
  ) |> 
  unnest(results) |> 
  mutate(city_state = fct_reorder(city_state, adj_OR))
```

* Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.
```{r}
plot_unsolved_sex = results_all  |> 
    arrange(adj_OR) |> 
  ggplot(aes(x = reorder(city_state, adj_OR), y = adj_OR, color = adj_OR)) +
    geom_point(size = 2) +
    scale_color_viridis_c(option = "turbo") +
    geom_errorbar(aes(ymin = `lower 95% CI`, ymax = `upper 95% CI`), size = 1, width = 0.2, alpha = 0.5) +
    labs(title = "Adjusted OR solved homicides by sex, by city", x = "City", y = "OR for Solved Homicides for male victims compared to females") +
    coord_flip() + 
    theme(
      panel.background = element_rect(fill = "grey95"), 
      axis.title = element_text(color = "black", size = 14), 
      axis.text = element_text(color = "black", size = 11), 
      title = element_text(color = "black", size = 18)
    ) +
  theme_bw()

plot_unsolved_sex

plot_save_path = "hw6_plots/OR_unsolved_sex_by_state.png"

ggsave(plot_save_path, plot_unsolved_sex, device = "png", width = 8, height = 9, units = "in", dpi = 300)
```

#### Problem 2

For this problem, we’ll use the Central Park weather data similar to data we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.
```{r, message = FALSE, warning=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. 

We’ll focus on a simple linear regression with `tmax` as the response with `tmin` and `prcp` as the predictors, and are interested in the distribution of two quantities estimated from these data:
  * r̂2
  * log(β̂1∗β̂2)

* Use 5000 bootstrap samples and, 
* for each bootstrap sample, produce estimates of these two quantities. 
```{r}
set.seed(123)

sim_estimates = weather_df |> 
   filter(!is.na(tmax) & !is.na(tmin) & !is.na(prcp)) |> 
  select(tmax, tmin, prcp
  ) |> 
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    rsquared = map(models, ~ summary(.x)$r.squared), 
    beta_product = map(models, ~ (coef(.x)["tmin"] * coef(.x)["prcp"]))
  ) |> 
    unnest(c(rsquared, beta_product)) |>
  mutate(
     logbeta = log(abs(beta_product)) 
     )|>
  select(.id, rsquared, beta_product, logbeta) 
  
```

* Plot the distribution of your estimates, and describe these in words. 
```{r, warning=FALSE, message=FALSE}
plot_sim_estimates_r2 = sim_estimates  |> 
  ggplot(aes(x = rsquared)) +
    geom_histogram(fill = "orange", color = "green3" ,alpha = 0.5) +
  labs(title = "R-squared values for simulated weather estimates", 
       subtitle = "tmax = tmin + prcp", 
       y = "Count", x = "R-squared value") +
    theme(
      panel.background = element_rect(fill = "grey95"), 
      axis.title = element_text(color = "black", size = 14), 
      axis.text = element_text(color = "black", size = 11), 
      title = element_text(color = "black", size = 18)
    ) +
  theme_bw()

plot_sim_estimates_r2

plot_save_path <- "hw6_plots/plot_sim_estimates_r2.png"

ggsave(plot_save_path, plot_sim_estimates_r2, device = "png", width = 6, height = 6, units = "in", dpi = 300)
```
The estimated R-squared values for the simulated weather data are approximately normal. The median of the data is at `r sim_estimates |> pull(rsquared) |> median()|> round(digits = 2)`. 

```{r, warning=FALSE, message=FALSE}
plot_sim_estimates_log = sim_estimates  |> 
  ggplot(aes(x = logbeta)) +
    geom_histogram(fill = "purple3", color = "pink" ,alpha = 0.5) +
  labs(title = "log(β̂1xβ̂2) values for simulated weather estimates", 
       subtitle = "tmax = tmin + prcp", 
       y = "Count", x = "log(β̂1xβ̂2)") +
    theme(
      panel.background = element_rect(fill = "grey95"), 
      axis.title = element_text(color = "black", size = 14), 
      axis.text = element_text(color = "black", size = 11), 
      title = element_text(color = "black", size = 18)
    ) +
  theme_bw()

plot_sim_estimates_log

plot_save_path <- "hw6_plots/plot_sim_estimates_log2.png"

ggsave(plot_save_path, plot_sim_estimates_log, device = "png", width = 6, height = 6, units = "in", dpi = 300)
```
The estimated R-squared values for the simulated weather data have a left skew. The median of the data is at `r sim_estimates |> pull(logbeta) |> median()|> round(digits = 2)`. 


* Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂2 and log(β̂0∗β̂1).  
  * Note: broom::glance() is helpful for extracting r̂2 from a fitted regression, and
  * broom::tidy() (with some additional wrangling) should help in computing log(β̂1∗β̂2).
```{r}
rsquared_logbeta_ci = sim_estimates |> 
  mutate(
    rsquared_lower = quantile(rsquared, 0.025, na.rm = TRUE),
    rsquared_upper = quantile(rsquared, 0.975, na.rm = TRUE), 
    logbeta_lower = quantile(logbeta, 0.0275, na.rm = TRUE), 
    logbeta_upper = quantile(logbeta, 0.975, na.rm = TRUE)
  )
```

#### Problem 3

In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

* Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
```{r}

```

* Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. 
* Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.
```{r}

```

* Compare your model to two others:
  * One using length at birth and gestational age as predictors (main effects only)
  * One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
* Make this comparison in terms of the cross-validated prediction error; 
* use crossv_mc and functions in purrr as appropriate.

```{r}

```

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.



